Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-03-08 WRITEUP/ms.tex
polished

\begin{abstract}
We study the dynamical properties of molecular clouds complexes (MC) in a prototypical galaxy (\flower) and its satellite galaxies at the Epoch of Reionization (EoR) using cosmological zoom-in simulation.
%
We identify MCs using H$_2$ density, and compare their masses, sizes, velocity dispersions, gas surface densities, and virial parameters ($\alpha_{\rm vir}$) with those observed in nearby and \z$\sim$\,2 galaxies.
%
The MCs span a mass range of $10^{6.5-9}$\,\Msun. Their velocity dispersions and gas surface densities are systematically higher than those observed in nearby galaxies, but comparable to starburst galaxies.
%
We compute the effective two-component Toomre $Q$ parameter (gas and stars) to examine regions that are unstable to local axisymmetric perturbations. 
%
We further perform virial analysis, which is accessible from observations, to examine regions likely 
to undergo collapse. Our analysis indicates that the MCs in the main disk of \flower are unbound, but its substructures have lower virial parameters and Toomre $Q$, which is consistent with the notion that collapsing structures occur within globally stable structures.
%
We find $\alpha_{\rm vir}\approx$\,2 for the MCs in the satellite galaxies, which we interpret as result of the weaker stellar feedback and less frequent \SF compared to \flower. Our results suggest that the more massive and bigger MCs compared to the Milky Way manifest from the higher gas mass fraction, surface density, and velocity dispersion.
% , which together sets the scale for fragmentation.
High resolution imaging of the first galaxies with the Atacama Large (sub-)Millimeter Array (ALMA) and the Next Generation Very Large Array (ngVLA) can furnish scrutinizing data to test our findings and shed light on \SF since the cosmic dark ages.
\end{abstract}
\keywords{methods: data analysis --
          galaxies: high-redshift --
          galaxies: ISM --
          galaxies: evolution --
          galaxies: formation --
          galaxies: starburst --
          stars: formation}

@@

The apparently large virial parameters seen in the MCs of \flower can be understood using the 
Toomre $Q$ parameter (see Figures~\ref{fig:Q}\,r$-$\,\ref{fig:Qeff}). 
We first note that fragmentation {\it can} happen in regions of low $Q$ (if we only consider 
instability against axisymmetric perturbations), but  
further evolution/collapse depends on the dynamics or equation of state of the gas. 
Such fragmentation is expected to take place at the
critical scale length $\lambda_{\rm crit} < 2 \pi^2 G \Sigma / \kappa^2$.
Second, this fragmentation scale is greater than the typical size of GMCs. This could be interpreted as a result of instability setting the scales for fragmentation, but the star-forming GMCs
correspond to the collapsing, denser, and cooler molecular structures that are on smaller scales. 
Thus, the high virial parameter found 
for most MCs in \flower, which is the ratio between the velocity dispersion (accounts for effect of pressure 
and rotation) and surface density, indicate that most of them are not collapsing structures. 
As such, some MCs lie in regions of $Q>1$. 
Only in (the denser) regions with a low Toomre $Q$ parameter and 
when energy can be dissipated on a timescale 
that is shorter than the timescale over which it gains energy (e.g., heating) 
correspond to star-forming {\it clumps} and {\it cores} with low virial parameters.
This would also (at least partially) explain the seemingly inconsistency between the high $\alpha_{\rm vir}$
nature of most MCs of \flower and its relatively high SFR.
That is, if most of the molecular gas in \flower are stable against collapse, 
how does it sustain its high SFR of $\sim$100\,\Msun\,yr\pmOne?
The discrepancy can be explained in two ways.
First, our results are limited by the resolution of the simulation, such that, in each MC, there
likely exist multiple smaller-scales molecular structures (e.g., clumps and cores),
as in the classical molecular cloud hierarchy. This re-iterates the statement about low $Q$ above.
Second, due to turbulence dissipation on small-scales, such (sub-)structures are no longer supported by
large-scale gravitational potential and differential rotation.
This in turn enables subregions of the MCs to collapse and potentially form OB associations out of these unbound
MCs \citep{Clark04a, Clark05a}.
Additional mechanisms causing non-axisymmetric perturbations (e.g., arms) to the gravitational potential
can also induce orbit crossing, shocks and dissipation in gas, further promoting turbulence dissipation
and collapse on smaller scales to trigger \SF.

@@

Virial analysis indicates that the MC/arms of the main disk of \flower are unbound, but the substructures have lower virial parameter.
This is consistent with the notion that collapsing structures result from gravitational instability within globally stable structures, which are
supported by turbulence and rotation on large scale.
We also find $\alpha_{\rm vir}\approx$\,2 for the MCs in the satellite galaxies, which we interpret as
result of the weaker stellar feedback compared to \flower 
MCs in the satellites are therefore likely collapsing structures.

We find no temporal variations in the MC dynamics over the course of 700\,Myr traced in our simulation, 
at least in terms of the scaling relations examined.
Our results are robust against the H$_2$ volume density threshold adopted (see Appendix \Sec{ncut}).

Determining the multi-phase ISM properties of early galaxies
is a critical piece to understanding the evolution and
assembly history of galaxies, since they set the pace
for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
Observations leveraging the combination of spatial-spectral imaging of
multi-band continuum and spectral line emission are crucial for better understanding
the role of \highz galaxy populations
in the context of galaxy evolution and the ISM physics behind their intense star formation in the early Universe.
Cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy
statistics and is subject to the sub-grid models adopted, 
serve as an useful avenue/tool for examining and making predictions on the morphology and dynamics of
the multi-phase ISM of the first galaxies.
%Future higher resolution zoom-in simulations would certainly be useful to understand
%the evolving dynamical properties of the lower-mass molecular structures in
%galaxies at EoR under the influence of gas inflow/outflow (i.e., accounting for energy injection at large scale due to infall/outflow).
These predictions can in turn be
tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby 
enabling us to test our findings and refine future simulations to shed light on the physical processes behind 
\SF since the cosmic dark ages.

>>>>

\begin{abstract}
We study the dynamical properties of molecular cloud complexes (MC) in a prototypical galaxy (\flower) and its satellite galaxies at the Epoch of Reionization (EoR) using cosmological zoom-in simulation.
%
We identify MCs using H$_2$ density, and compare their masses, sizes, velocity dispersions, gas surface densities, and virial parameters ($\alpha_{\rm vir}$) with those observed in nearby and \z$\sim$\,2 galaxies.
%
The MCs span a mass range of $10^{6.5-9}$\,\Msun. Their velocity dispersions and gas surface densities are systematically higher than those observed in nearby galaxies, but comparable to starburst galaxies.
%
We compute the effective two-component Toomre $Q$ parameter (gas and stars) to examine regions that are 
gravitationally unstable.
%
We further perform virial analysis, which is accessible from observations, to examine regions likely 
to undergo collapse. Our analysis indicates that the MCs in the main disk of \flower are unbound, but its substructures have lower virial parameters and Toomre $Q$, which is consistent with the notion that collapsing structures occur within globally stable structures.
%
We find $\alpha_{\rm vir}\approx$\,2 for the MCs in the satellite galaxies, which we interpret as result of the weaker stellar feedback 
compared to \flower. Our results suggest that the more massive and bigger MCs compared to the Milky Way manifest from the 
higher gas mass fraction, surface density, and velocity dispersion.
% , which together sets the scale for fragmentation.
High resolution imaging of the first galaxies with the Atacama Large (sub-)Millimeter Array (ALMA) and the Next Generation Very Large Array (ngVLA) can furnish scrutinizing data to test our findings and shed light on \SF since the cosmic dark ages.
\end{abstract}
\keywords{methods: data analysis --
          galaxies: high-redshift --
          galaxies: ISM --
          galaxies: evolution --
          galaxies: formation --
          galaxies: starburst --
          stars: formation}

@@

The apparently large virial parameters seen in the MCs of \flower can be understood using the 
Toomre $Q$ parameter (see Figures~\ref{fig:Q}\,$-$\,\ref{fig:Qeff}). 
We first note that fragmentation {\it can} happen in regions of low $Q$ (if we only consider 
instability against axisymmetric perturbations), but  
further evolution/collapse depends on the dynamics or equation of state of the gas. 
Such fragmentation is expected to take place at the
critical scale length $\lambda_{\rm crit} < 2 \pi^2 G \Sigma / \kappa^2$.
Second, this fragmentation scale is greater than the typical size of GMCs. This could be interpreted as a result of instability setting the scales for fragmentation, but the star-forming GMCs
correspond to the collapsing, denser, and cooler molecular structures that are on smaller scales. 
Thus, the high virial parameter found 
for most MCs in \flower, which is the ratio between the velocity dispersion (accounts for effect of pressure 
and rotation) and surface density, indicate that most of them are not collapsing structures. 
As such, some MCs are found to lie in regions of $Q>1$. 
% some MCs at Q < 1, some are marginally stable against m=0 mode perturb, some are Q>1.
Only in (the denser) regions with a low Toomre $Q$ parameter and 
when energy can be dissipated on a timescale 
that is shorter than the timescale over which it gains energy (e.g., heating) 
correspond to star-forming {\it clumps} and {\it cores} with low virial parameters.
This would also (at least partially) explain the seemingly inconsistency between the high $\alpha_{\rm vir}$
nature of most MCs of \flower and its relatively high SFR.
That is, if most of the molecular gas in \flower are stable against collapse, 
how does it sustain its high SFR of $\sim$100\,\Msun\,yr\pmOne?
The discrepancy can be explained in two ways.
First, our results are limited by the resolution of the simulation, such that, in each MC, there
likely exist multiple smaller-scales molecular structures (e.g., clumps and cores),
as in the classical molecular cloud hierarchy. This re-iterates the statement about low $Q$ above.
Second, due to turbulence dissipation on small-scales, such (sub-)structures are no longer supported by
large-scale gravitational potential and differential rotation.
This in turn enables subregions of the MCs to collapse and potentially form OB associations out of these unbound
MCs \citep{Clark04a, Clark05a}.
Additionally, mechanisms causing non-axisymmetric perturbations (e.g., arms) to the gravitational potential
can induce orbit crossing, shocks and dissipation in gas, promoting turbulence dissipation
and collapse on smaller scales to trigger \SF.

@@

Virial analysis indicates that MC/arms of the main disk of \flower are unbound, but substructures have
lower virial parameter. Some of these sub-MCs have Toomre $Q_{\rm eff}$ parameters consistent with being 
gravitationally unstable. This is consistent with the notion that collapsing structures resulting from
gravitational instability occur within globally stable structures, which are
supported by turbulence and rotation on large scale.
We also find $\alpha_{\rm vir}\approx$\,2 for MCs in the satellite galaxies, which we interpret as
result of the weaker stellar feedback compared to \flower.
MCs in the satellites are therefore likely collapsing structures.

In terms of the scaling relations examined, we do not observe any significant
temporal variations in the MC dynamics over the course of 700\,Myr traced in our simulation.
Similarly, we do not find any significant dependence on the H$_2$ volume density threshold adopted
in the scaling relations examined (see Appendix \Sec{ncut}). 
Our results are thus robust against the density threshold adopted.

Determining the multi-phase ISM properties of early galaxies
is a critical piece to understanding the evolution and
assembly history of galaxies, since they set the pace
for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
Observations leveraging the combination of spatial-spectral imaging of
multi-band continuum and spectral line emission are crucial for better understanding
the role of \highz galaxy populations
in the context of galaxy evolution and the ISM physics behind their intense star formation in the early Universe.
Cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy
statistics and is subject to the sub-grid models adopted, 
serve as an useful avenue/tool for examining and making predictions on the morphology and dynamics of
the molecular ISM of the first galaxies.
%Future higher resolution zoom-in simulations would certainly be useful to understand
%the evolving dynamical properties of the lower-mass molecular structures in
%galaxies at EoR under the influence of gas inflow/outflow (i.e., accounting for energy injection at large scale due to infall/outflow).
These predictions can in turn be
tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby 
enabling us to test our findings and refine future simulations to shed light on the physical processes behind 
\SF since the cosmic dark ages.
